Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-07-08 edubots_surveys.Rnw
Revising the new Introduction

% TODO: introduction
The introduction of new technologies in the classroom involves a combination of teacher training, student enthusiasm and possibly acquiring new skills by both. Some technologies are readily adapted, but others require longer time for full adoption. In most cases, the magnitude of the uptake of the new technology impinges on the collaboration of all the parties using it. Hence, examining the users’ attitudes and the opinions is an important first step in boosting this collaboration and reaping the benefits of the technology.  Recently, a surge in the use of synchronous instant messaging tools in higher education such as WhatsApp, Telegram and Facebook Messenger took place. Some messaging applications have built in chatbots that are able to support a synchronous conversation between the different parties \cite{differ.chat}, while other chatbots are developed as standalone systems and can possibly be attached to a messaging application using an API (Application Program Interface).

A chatbot is a program, sometimes developed using Artificial Intelligence, techniques, that allows an application to communicate in a similar way as humans do \cite{Gong2008}, using, in some cases, natural language. Indeed, most of them would be, apparently, close to passing a classic Turing test \cite{moor2003turing}, since they are able to answer almost any question fluently, and even pose questions in their conversations. The recent advancement in this discipline has played an important role in many fields especially in education and online tutoring \cite{clarizia2018,Smutny2020}. These automatic systems facilitate the delivery of personalized learning by adapting to students’ pace of learning and providing customized online tutoring outside the classroom. Thus, chatbots can significantly contribute to providing interactive learning experiences as well as improving individual attention \cite{agarwal2020}.
 In this sense, chatbot technology offers a great opportunity for the improvement of tutoring systems \cite{Daniel2016,agarwal2020}, as not all students are comfortable with face-to-face tutoring with the instructor. In fact, in many cases, some students experience stress because of their need to ask a question in front of the entire class. This is why many students refrain to silence in the classroom and tend to contact their teachers later via email. Consequently, this can lead to not only a delay in obtaining answers to their questions but also can cause a significant increase in their teachers’ workload, especially when factoring in the disproportion between the class size and the number of teachers. Hence, the chatbot technology has a potential to mitigate this problem by providing answers to students’ questions and facilitating a dynamic and autonomous learning experience \cite{Griol2014,kim2020}. In addition, using an automated system such as a chatbot, can draw teachers’ attention to topics that students struggle to comprehend or need further assistance in understanding.

Given the potential benefits that chatbots could bring to the classroom, the focus of the EDUBOTS KA2 European project \footnote{https://www.edubots.eu}, the sponsor of this study, is to gather, analyse and disseminate best practice of the use of chatbots in higher education institutions.  The study presented in this paper involves two phases of data collection.\begin{itemize}
\item The first phase aimed to collect feedback from students about their actual needs \cite{MoraChatbots2021}.  Hence, the objective was to collect information about students’ preferences of tools/applications for chatting and messaging, how they use them in educational context, who do they like to be with in the class messaging groups, and their expectations of the chatbots in terms of assisting them during their learning process. Therefore, two surveys for bachelor and master degree students at the University of Granada (Spain) were designed and answers from more than 250 students were collected.   The key findings of the students surveys reveal that they prefer to use messaging applications, such as Telegram or WhatsApp as they are familiar to them. Students also expressed interest in chatbots that can assist them with organizing their course schedules, help them access their assignment grades and facilitate searching for course’s resources. 
\item The results of the first phase were fed into the second phase of the study, which focused on teachers’ opinions about using chatbots and messaging applications in classes.
\end{itemize}

Another survey was developed for this purpose. Responses from 300 teachers were collected and analysed.  This study focuses on discussing the implications and the impact of the development of future technology-enhanced tools and educational institutions’ policies by finding answers to the following research questions:
\begin{itemize}
\item RQ1 - Are teachers already using messaging apps in their classes? 
\item RQ2 - Which kind of chatbots would teachers find useful in their classes? 
\item RQ3 - Which kind of interaction do teachers prefer with their students?
\item RQ4 - What kind of interaction media features do teachers value the most?
\end{itemize}

The remainder of the paper is organized as follows: Next is an overview of what current research has found about the use of messaging applications, including chatbots, in the classroom. The methodology used in the survey is presented in Section \ref{sec:meth}, and the results of the survey are presented next in Section \ref{sec:res}. Finally, we discuss these results and conclude with a series of recommendations in the section that closes the paper.

>>>>

% TODO: introduction
The introduction of new technologies in the classroom involves a combination of teacher training, student enthusiasm and possibly acquiring new skills by both. Some technologies are readily adapted, but others require longer time for full adoption. In most cases, the magnitude of the uptake of the new technology impinges on the collaboration of all the parties using it. Hence, examining the users’ attitudes and their opinions is an important first step in boosting this collaboration and reaping the benefits of the technology.  Recently, a surge in the use of synchronous instant messaging tools in higher education such as WhatsApp, Telegram and Facebook Messenger took place. Some messaging applications have built in chatbots that are able to support a synchronous conversation between the different parties \cite{differ.chat}, while other chatbots are developed as standalone systems and can possibly be attached to a messaging application using an API (Application Program Interface).

A \textit{chatbot} is a program, sometimes developed using Artificial Intelligence techniques, able to communicate in a similar way as humans do \cite{Gong2008}, using, in some cases, natural language. Indeed, most of these applications would be, apparently, close to passing a classic Turing test \cite{moor2003turing}, since they are able to answer almost any question fluently, and even pose own questions in their conversations. The recent advancement in this discipline has played an important role in many fields especially in education and online tutoring \cite{clarizia2018,Smutny2020}. These automatic systems facilitate the delivery of personalized learning by adapting to students’ pace of learning and providing customized online tutoring outside the classroom. Thus, chatbots can significantly contribute to providing interactive learning experiences as well as improving individual attention \cite{agarwal2020}.
 In this sense, chatbot technology offers a great opportunity for the improvement of tutoring systems \cite{Daniel2016,agarwal2020}, as not all students are comfortable with face-to-face tutoring with the instructor. In fact, in many cases, some students experience stress because of their need to ask a question in front of the entire class. This is why many students refrain to speak  
 % Antonio - I guess refrain to silence is the opposite of what we wanted to say, isn't it? :D
 in the classroom and tend to contact their teachers later via email.
 Consequently, this can lead to not only a delay in obtaining answers to their questions but also can cause a significant increase in their teachers’ workload, especially when factoring in the disproportion between the class size and the number of teachers. Hence, the chatbot technology has a potential to mitigate this problem by providing answers to students’ questions and facilitating a dynamic and autonomous learning experience \cite{Griol2014,kim2020}. In addition, using an automated system such as a chatbot, can draw teachers’ attention to topics that students struggle to comprehend or need further assistance in understanding.

Given the potential benefits that chatbots could bring to the classroom, the focus of the EDUBOTS KA2 European project \footnote{https://www.edubots.eu}, the sponsor of this study, is to gather, analyse and disseminate best practices on the use of chatbots in higher education institutions.  
% Antonio - Actually there was a previous study focused on students and the present one, focused on teachers. I'll try to explain this, please revise the text below. ;)
With this aim, the authors conducted a previous study aimed to collect feedback from students about their actual needs \cite{MoraChatbots2021}.  Hence, the objective was to collect information about students’ preferences of tools/applications for chatting and messaging, how they use them in educational context, who do they like to be with in the class messaging groups, and their expectations of the chatbots in terms of assisting them during their learning process. Therefore, two surveys for bachelor and master degree students at the University of Granada (Spain) were designed and answers from more than 250 students were collected. The key findings of the students surveys reveal that they prefer to use messaging applications, such as Telegram or WhatsApp as they are familiar to them. Students also expressed interest in chatbots that can assist them with organizing their course schedules, help them access their assignment grades and facilitate searching for course’s resources.

The results of this first phase were fed into the second phase of the study, which is addressed in the present work. It is focused on the other main actor in education process: the teacher. Thus, we have now collected and analised teachers’ opinions about using chatbots and messaging applications in classes.
To this end two new surveys have been developed, and responses from almost 500 teachers (282 + 205) were collected and analysed.  This study focuses on discussing the implications and the impact of the development of future technology-enhanced tools and educational institutions’ policies by finding answers to the following research questions:
\begin{itemize}
\item RQ1 - Are teachers already using messaging apps in their classes? 
\item RQ2 - Which kind of chatbots would teachers find useful in their classes? 
\item RQ3 - Which kind of interaction do teachers prefer with their students?
\item RQ4 - What kind of interaction media features do teachers value the most?
\end{itemize}

We will try to answer these questions through the analysis of questions and answers in the two surveys. Eventually, what we want to obtain is a series of recommendations to make a successful deployment of chatbot technologies (and, in some cases, general instant messaging applications) in higher education.

The remainder of the paper is organized as follows: Next is an overview of what current research has found about the use of messaging applications, including chatbots, in the classroom. The methodology used in the surveys is presented in Section \ref{sec:meth}, and the results of the surveys are presented next in Section \ref{sec:res}. Finally, we discuss these results and conclude with a series of recommendations in the section that closes the paper.
